Continuous gravitational waves in the lab: recovering audio signals with a table-top optical microphone

### 2020-06-16 paper/ifo-appendix.tex
going through a few more comments from Andrew

\section{Open-source code}
\label{app:code}
This project was implemented in python3~\cite{python} scripts as well as jupyter notebooks~\cite{jupyter}~\cite{ipython} and used the packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.

The current build and sample recordings of the optical microphone can be found at:
\url{https://github.com/daccordeon/gravexplain}
% \begin{verbatim}
% README
% \end{verbatim}

\section{A common mistake for those new to wandering frequency signals}
\label{app:phase_gotcha}

We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$. A common mistake is to appeal to the standard form $\sin{(2 \pi f t)}$, where f is a constant frequency, and guess $\sin{(2 \pi f(t) t)}$. However, this fails to address the phase accumulated earlier in the signal. Instead, the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$. Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
% This formula is also useful when performing frequency modulation.

Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram, it is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}. A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}. This design was based off standard examples of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.

\begin{figure*}
	\includegraphics[width=.65\textwidth]{figures/circuit_diagram_2.pdf}
	\caption{Circuit diagram for photodiode reading. Photodiode in reverse-bias over an op-amp, analog signal then passed through Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC). The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.}
	\label{fig:circuit_diagram}
\end{figure*}

\begin{figure*}%[H]
	\includegraphics[width=0.5\textwidth]{figures/circuit_pic3.pdf}
	\caption{Photo of photodiode circuit assembled on breadboard.}
	\label{fig:circuit_pic3}
\end{figure*}

\section{Interferometric intensity relation derivation}
\label{app:intensity_derivation}

@@

>>>>

\section{Open-source code}
\label{app:code}
This project is implemented in python3~\cite{python} scripts as well as jupyter notebooks~\cite{jupyter}~\cite{ipython} and used the packages of numpy~\cite{numpy}, scipy~\cite{scipy}, matplotlib~\cite{matplotlib}, tqdm~\cite{tqdm}, and logmmse~\cite{logmmse}.
The current build and sample data can be found at:
\url{https://github.com/daccordeon/gravexplain}
% \begin{verbatim}
% README
% \end{verbatim}

\section{Interferometric relation between path length and optical intensity}
\label{app:intensity_derivation}

@@

%\section{A common mistake for those new to wandering frequency signals}
%\label{app:phase_gotcha}
%We want to generate a sine wave that changes frequency over time so that the frequency at some time is given by a function $f(t)$. A common mistake is to appeal to the standard form $\sin{(2 \pi f t)}$, where f is a constant frequency, and guess $\sin{(2 \pi f(t) t)}$. However, this fails to address the phase accumulated earlier in the signal. Instead, the standard form in fact says that for some $\sin{(F(t))}$ it is true that $\frac{dF}{dt}(t) = 2 \pi f(t)$, where $f(t)$ gives the frequency of $\sin{(F(t))}$ at time $t$. Therefore, the correct form for a wandering tone is $\sin{(\int{2 \pi f(t) dt})}$.
%% This formula is also useful when performing frequency modulation.

Figure~\ref{fig:circuit_diagram} shows the photodiode circuit diagram. 
It is also available at \url{https://crcit.net/c/e397dcc2166943d69155f9dac1e27bce}. A photo of the breadboard and Raspberry Pi pins is shown in Figure~\ref{fig:circuit_pic3}. This design was based off standard examples of photo-detectors, connecting an ADC to a Pi, and Sallen-Key filters.

\begin{figure*}
	\includegraphics[width=.65\textwidth]{figures/circuit_diagram_2.pdf}
	\caption{\label{fig:circuit_diagram}
Circuit diagram for reading the photodiode. 
The photodiode is connected in reverse-bias across an op-amp. 
The analog op-amp output is passed through a Sallen-Key anti-aliasing filter tuned to 16kHz, then into an analog-to-digital converter (ADC). 
The digitised signal is then read by the special purpose input (SPI) pins of a Raspberry Pi in standard configuration.
}
\end{figure*}

\begin{figure*}%[H]
	\includegraphics[width=0.5\textwidth]{figures/circuit_pic3.pdf}
	\caption{\label{fig:circuit_pic3}
Photodiode circuit assembled on breadboard.
}
\end{figure*}

### 2020-06-16 paper/ifo-complexAudio.tex
going through a few more comments from Andrew

% Optical microphone
% aim of optical microphone: play back injected sounds

A natural succession to observing simple tones is to use complex audio signal, such as music and speech. 
The Viterbi algorithm and tracking more broadly is not applicable to these signals---music and speech are constantly changing over a broad spectrum. 
Instead of tracking a changing frequency over time, here we are interested in being able to record the interference pattern and play back these complex audio signals. 
We use the interferometer as an `optical microphone', using light to capture sound. 
The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.

This section is indirectly related to the advanced signal processing required to extract gravitational-wave signals. 
The optical microphone provides an interesting independent demonstration for a broader physics and engineering audience, particularly for the undergraduate lab.

We describe additional components to the demonstration in Section~\ref{sec:photodiode} and our initial results in Section~\ref{sec:initialResultsOpMic}.
Filter and speech enhancement techniques applied to the data are detailed in Sections~\ref{sec:simple_filters} and~\ref{sec:logmmse} respectively.

\subsection{Recording data via photodiode}
\label{sec:photodiode}
% advanced method: explain how we capture data

@@

A photodiode is an electrical component that acts as a regular diode wheen no light is incident on it, blocking any current flow in the `reverse' direction.
As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction. 
% Fundamentally, a current is created in the photodiode by the photo-electric effect.
We place a (OSRAM BPW21) photodiod in reverse-bias over an (LM358) op-amp. 
This creates a photo-detector that produces a voltage dependent on the incident intensity. 
The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} (however we see no difference when monitoring at other positions in the interference pattern).
The photodiode was mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.

The voltage signal from the photo-detector is captured by a (MCP3008) $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2. 
Together, these sampled the signal at roughly $16\,{\rm kHz}$.

Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing into the detected range. 
To prevent this, we include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC. 
This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
We also place another cloth screen over the face of the photodiode, as the reading without it, after gaining through the op-amps, was too high for the ADC.
% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).

@@

\subsection{Initial results of the optical microphone}
\label{sec:initialResultsOpMic}
% breakdown of all the filters used along the way

The optical microphone is tested on varied recordings of speech from different people and on music ranging from simple melodies to full track songs. 
Each recording is around one minute long with processing applied to ten second segments. 
For each observation, the audio is played through the speaker whilst the Pi script records the signal from the photodiode. 
During recordings, care is taken to ensure minimal activity around the interferometer to reduce environmental noise. 
The timeseries is then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python} as before.

We find that initial recordings are extremely noisy with a loud bass hum throughout.
Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond the $8$th harmonic thereof.
We find that the mains power signal is present in data recorded without an injected audio signal as shown by the power spectral density (effectively the norm squared of the Fourier spectrum) in Figure~\ref{fig:psd_noise}.
The mains power is also present in data taken with the photodiode in complete darkness with a similar but weaker distribution as that shown in Figure~\ref{fig:psd_noise}).
The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination.

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
	\caption{Power spectral density (PSD) of the optical microphone system with no injected audio (i.e.\ with the speaker off). We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the surrounding room’s lighting and even cooling), but otherwise is a fairly white spectrum.}
	% except for the broad, shallow maximum at 750Hz}
	\label{fig:psd_noise}
\end{figure}

\subsection{Applying filters to the data}
\label{sec:simple_filters}

Here we describe several approaches to remove the $50\,{\rm Hz}$ mains power signal and its harmonics.

Simply zeroing the frequency channel of each $50\,{\rm Hz}$ harmonic in the Fourier spectrum is unsuccessful. 
Effectively, this multiplies the spectrum by a rectangular comb filter. 
It does remove the mains harmonics, however it audibly ruins the rest of the signal. 
Applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter. 
The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function, which significantly damages the signal.

@@

Applying a high-pass filter with cut-off frequency $150\,{\rm Hz}$ to the signal works well at removing the $50\,{\rm Hz}$ and $100\,{\rm Hz}$ harmonics.
However, mains harmonics $> 100\,{\rm Hz}$ and remain after filtering. 
The region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}. 
Therefore, although a high-pass filter with a higher cut-off removes more harmonics, the resultant signal is unrecognisable. 
Applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.

@@

Band-passing the signal similarly does not address the remaining harmonics above $100\,{\rm Hz}$. 
However, it does low-pass away other noise at higher frequencies. 
The best performing band-pass filter we found was a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}). 
For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\epsilon$, the frequency response (attenuation at each frequency) is,
\begin{equation}
\label{eq:butterworth}
H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\frac{\omega}{\omega_c})^{2n}}},
\end{equation}
as a function of frequency $\omega$. 
To construct a Butterworth band-pass filter, the above low-pass filter is combined with a high-pass filter of similar form. 
The response of the filter used is shown in Figure~\ref{fig:butterworth}. 
This filter improves on the above simpler filters but recordings of speech still have very low intelligibility, they can be recognised as speech but not understood.

The ideal filter would be one that i.) completely attenuates the undesired parts of the spectrum, ii.) does not change the rest of the spectrum, and iii.) has smooth edges so as to not damage the signal through convolution. 
These three conditions are contradictory and any filter must compromise between them. 
The above simple filters fail at one or more of each of these conditions. 
The combs ruin the signal and the high-passes do not remove the noise. 
The Butterworth filter performs the best audibly as it is optimised to be `maximally flat' within the band region, prioritising the second condition.

@@

\subsection{The logMMSE estimator}
\label{sec:logmmse}
% logMMSE

Speech enhancement for a noisy channel is a classic problem in signal processing. 
Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods. 
They find the best performing statistical technique considered to be a logMMSE (alt.\ MMSE-LSA) estimator. 
This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
Historically this technique minimises with respect to the spectral amplitude~\cite{EphraimSpeechEU_MMSE}. 
However in speech processing, use of the logarithm seems to produce better results as percieved by the human ear~\cite{SubjectiveComparison}.

In this work we used an existing implementation~\cite{logmmse} of the logMMSE estimator. 
Applying this estimator to the optical microphone recordings produced dramatically cleaner results.

In Figures~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} we present results from playing a speech recording through the optical microphone. 
Figure~\ref{fig:logMMSE_timeseries} shows the time series from before (top panel) and after (bottom panel) applying the logMMSE filter. 
Figure~\ref{fig:logMMSE_spectrum} shows the corresponding spectra before (top panel) and after (bottom panel) the logMMSE filter. 
Comparing the before and after spectra shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.

@@

We also find improvement with music recordings.
After applying the estimator, simple chords and drums can be heard clearly. 
However more complex melodies are sometimes completely absent from the resulting time series. 
From observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
This could be a perceptual effect of listening to the lossy recording, but seems more likely to be a frequency dependence somewhere in the optical microphone system. 
Speculating, perhaps the amplitude of speaker-mirror oscillations is higher at low frequencies and so instruments like electric bass and drums are louder in the results.

Overall, we find that the intelligibility of speech signals remains low. 
The voice sounds mumbled and indistinct. 
The estimator removes most of the background noise, but does not clarify the speech. 
To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all. 
This requires a better understanding of the system and further improvements to the experiment if true.
% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio. However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference. An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.

>>>>

% Optical microphone
% aim of optical microphone: play back injected sounds
A natural successor to simple tomes is complex audio, such as music and speech. 
Complex audio signals are not quasimonochromatic, unlike continuous gravitational wave signals, so the Viterbi algorithm as implemented in Section~\ref{} is not applicable directly.
Instead, in this section, we explore how to use the interferometer as an `optical microphone' to detect and play back complex audio signals, using light to capture sound. 
In doing so, we experiment with a hierarchy of passive filters which suppress noise yet make no assumption about the specific form of the signal, unlike the Fourier-based maximum likelihood matched filter tuned to sinusoidal signals in Section~\ref{} and Appendix~\ref{}. 
The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.
It is also analogous to the detection of unmodelled gravitational-wave signals, e.g. broadband bursts from supernovae~\cite{}. 
%It is also analogous to the detection of unmodelled gravitational wave signals, e.g. broadband bursts from supernovae. [refs; my last sentence can be improved]

We describe additional components to the demonstration in Section~\ref{sec:photodiode} and our initial results in Section~\ref{sec:initialResultsOpMic}.
Filter and speech enhancement techniques applied to the data are detailed in Sections~\ref{sec:simple_filters} and~\ref{sec:logmmse} respectively.
This section is indirectly related to the advanced signal processing required to extract gravitational-wave signals. 
The optical microphone provides an interesting independent demonstration for a broader physics and engineering audience, particularly for the undergraduate lab.

\subsection{Photodiode}
\label{sec:photodiode}
% advanced method: explain how we capture data

@@

A photodiode is an electrical component that acts as a regular diode when no light is incident on it, blocking any current flow in the `reverse' direction.
As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction. 
% Fundamentally, a current is created in the photodiode by the photo-electric effect.
We place a OSRAM BPW21 photodiode in reverse-bias over an LM358 op-amp. 
This creates a photo-detector that produces a voltage dependent on the incident intensity. 
The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering}.
We see no difference when monitoring at other positions in the interference pattern.
The photodiode is mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.

The voltage signal from the photo-detector is captured by a MCP3008 $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2. 
Together, these sample the signal at roughly $16\,{\rm kHz}$.

Sampling any frequency component of the analog signal above the Nyquist frequency of $8\,{\rm kHz}$ leads to aliasing into the detected range. 
To prevent this, we include an anti-aliasing Sallen-Key filter~\cite{sallen_key_filter} tuned to $16\,{\rm kHz}$ before the ADC. 
This component attenuates any frequencies above $8\,{\rm kHz}$, before they are digitally sampled.
We also place another cloth screen over the face of the photodiode, as the reading without it is too high for the ADC after gaining through the op-amps. 
% It’s not clear what limited the sample rate to 16kHz, but it was likely non-optimal reading of the ADC by the Pi script.
% The ADC used (MCP3008) is quoted at 200kHz (or ksps, kilo samples per second).

@@

\subsection{Anti-aliased output}
\label{sec:initialResultsOpMic}
% breakdown of all the filters used along the way

The optical microphone is tested on recordings of speech from different people and on music ranging from simple melodies to full track songs. 
Each recording is around one minute long with processing applied to ten second segments. 
For each observation, the audio is played through the speaker whilst the Pi script records the signal from the photodiode. 
During recordings, care is taken to ensure minimal activity around the interferometer to reduce environmental noise. 
The timeseries is then directly converted to a .wav file and played as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python} as before.

We find that initial recordings are extremely noisy with a loud bass hum throughout.
Each $10\,{\rm s}$ Fourier spectrum shows dominant AC mains noise with power ranging from the fundamental of the $50\,{\rm Hz}$ mains power up to and beyond its $8$th harmonic.
We find that the mains power signal is present in data recorded without an injected audio signal as seen in the power spectral density in Figure~\ref{fig:psd_noise}.
The mains power is also present in data taken with the photodiode in complete darkness with a similar but weaker distribution as that shown in Figure~\ref{fig:psd_noise}).
The origin of the mains power signal is likely to be from the photodiode signal, or environmental noise from the rooms lighting and air conditioning, or a combination.

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/podo_noise_psd_zoom-cropped.pdf}
	\caption{\label{fig:psd_noise}
Power spectral density (PSD) of the optical microphone  with no injected audio (i.e.\ with the speaker off). 
We see strong power from the $50\,{\rm Hz}$ mains hum and its harmonics (most likely from the photodiode circuit and the room’s lighting and cooling). 
Otherwise the PSD is fairly white. 
}
\end{figure}

\subsection{Additional filters}
\label{sec:simple_filters}

Here we describe several approaches to remove the $50\,{\rm Hz}$ mains hum and its harmonics.

Simply zeroing the frequency channel of each $50\,{\rm Hz}$ harmonic in the Fourier spectrum is unsuccessful. 
Effectively, this multiplies the spectrum by a rectangular comb filter. 
It does remove the mains harmonics, however it audibly ruins the rest of the signal. 
Applying a filter in frequency space is equivalent to convolving the signal with the inverse Fourier transform of that filter. 
The inverse Fourier transform of a rectangular comb filter (a set of boxcars) is a complicated sinc-like function, which significantly corrupts the signal.

@@

Applying a high-pass filter with cut-off frequency $150\,{\rm Hz}$ to the signal works well at removing the $50\,{\rm Hz}$ and $100\,{\rm Hz}$ harmonics.
However, mains harmonics above $100\,{\rm Hz}$ remain. 
The region above $100\,{\rm Hz}$ carries a lot of the fundamental frequencies of speech and music~\cite{speech_intelligibility}. 
Therefore, although a high-pass filter with a higher cut-off removes more harmonics, the played-back signal is unrecognisable. 
Applying a high-pass filter to the logarithm of the signal spectrum does not significantly improve on the above simple high-pass filter.

@@

Band-passing the signal similarly does not address the remaining harmonics above $100\,{\rm Hz}$. 
However, it does suppress other noise at higher frequencies. 
The best performing band-pass filter we find is a fifth order ($n = 5$) Butterworth filter over the range $150\,{\rm Hz}$--$3\,{\rm kHz}$ (beyond $2\,{\rm kHz}$ speech information content drops significantly~\cite{speech_intelligibility}). 
For a low-pass Butterworth filter with cut-off frequency $\omega_c$ and gain $\epsilon$, the frequency response (attenuation at each frequency) is,
\begin{equation}
\label{eq:butterworth}
H(\omega) = \frac{1}{\sqrt{1+\epsilon^2 (\frac{\omega}{\omega_c})^{2n}}},
\end{equation}
as a function of frequency $\omega$. 
To construct a Butterworth band-pass filter, the above low-pass filter is combined with a high-pass filter of similar form. 
The response of the filter used is shown in Figure~\ref{fig:butterworth}. 
This filter improves on the above simpler filters but recordings of speech still have low intelligibility.
They can be recognised as speech without being understood.

The ideal filter would be one that i) completely attenuates the undesired parts of the spectrum, ii) does not change the rest of the spectrum, and iii) has smooth edges so as to not damage the signal after convolution. 
These three conditions are contradictory and any filter must compromise between them. 
The above simple filters fail one or more of i)--iii). 
The Butterworth filter performs the best as it is optimised to be `maximally flat' within the band region, prioritising the second condition.

@@

\subsection{logMMSE estimator}
\label{sec:logmmse}
% logMMSE

Speech enhancement for a noisy channel is a classic problem in signal processing. 
Hu~\&~Loizou~(2006)~\cite{SubjectiveComparison} compare 13 speech enhancement methods and find the best to be a log minimum mean-square error (log-MMSE) estimator. 
This type of estimator is based on work by Ephraim~\&~Malah~(1984)~\cite{Ephraim1984SpeechEU_logMMSE} into speech enhancement by minimising the mean square error (MSE) to the logarithm of the spectral (Fourier) amplitude.
The logarithm approximates the response of the human ear~\cite{SubjectiveComparison}.

In this work we use an existing implementation~\cite{logmmse} of the logMMSE estimator. 
Applying this estimator to the optical microphone output produced dramatically cleaner results.

In Figures~\ref{fig:logMMSE_timeseries} and~\ref{fig:logMMSE_spectrum} we present results from playing a speech recording through the optical microphone. 
Figure~\ref{fig:logMMSE_timeseries} shows the anti-aliased photodiode time series from before (top panel) and after (bottom panel) applying the logMMSE filter. 
Figure~\ref{fig:logMMSE_spectrum} shows the corresponding spectra before (top panel) and after (bottom panel) the logMMSE filter. 
Comparing the before and after spectra shows significant attenuation of all mains harmonics as well as a general smoothing of the spectrum.

@@

We also find improvement with music recordings.
After applying the estimator, simple chords and drums can be heard clearly. 
However more complex melodies are sometimes completely absent. 
From observation, this is especially true for instruments of certain timbres (harmonic profiles), in particular flutes and violins.
This could be a perceptual effect or a frequency dependence somewhere in the optical microphone. 
Speculating, perhaps the speaker-mirror oscillates more at at low frequencies and so instruments like electric bass and drums are louder in the results.

Overall, we find that the intelligibility of speech remains low. 
The played-back voice sounds mumbled and indistinct. 
The estimator removes most of the background noise, but does not clarify the speech. 
To address these problems with the recordings we need to determine whether the signals that are audibly missing (the diction in the speech and complex melodies in music) are indeed being transmitted through the optical microphone at all. 
This requires a better understanding of the system.
% The fringe counting problem of the detector, wherein multiple fringes passing over the array in a single speaker deflection artificially raises the frequency, was noticeable in the speech recordings when compared to the source audio. However, upon comparing injected pure tones, the increase appears to be minimal and inconsistent, at around 5Hz maximum difference. An simple explanation is that the speaker deflections are small compared to manual pressing on the optical table.

### 2020-06-16 paper/ifo-conclusions.tex
going through a few more comments from Andrew

In this work we use a table-top Michelson interferometer as an analogue to a gravitational-wave detector. 
Signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirror. 
We consider monochromatic and slowly-wandering frequency signals as an analogue to continuous wave signals as well as more complex audio recordings of speech and music. 
% This serves to demonstrate signal processing techniques.

In Sections~\ref{sec:ifo} and ~\ref{sec:single_tone} we demonstrate the recovery of a monochromatic and wandering frequency signal. 
A webcam is used to track the interference pattern and the injected signal is recovered succesffully with the Viterbi algorithm. 
These sections direcly demonstrate the continuous wave search techniques used in gravitational-wave data analysis. 
%Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer. We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.

Substituting the webcam for the increased frequency response of a photodiode circuit read by a Raspberry Pi, we turned the interferometer into an optical microphone.
We demonstrate that the optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech. 
Dominant $50\,{\rm Hz}$ mains hum in the recordings led to an exploration of speech enhancement techniques to remove it, including the well performing logMMSE estimate. 
Further avenues of exploration for this work include changes to the demonstration hardware to further remove the mains harmonics as well as investigating a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.

The demonstration presented here has good potential to be adapted for use in the undergraduate laboratory. 
It can be used to demonstrate topics of interest to astrophysics and digital communications including: interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement. 
%These demonstrations have good potential to be adapted into the undergraduate laboratory. There is much to explore in the physics of the interferometer and the astronomy of the motivating continuous gravitational wave searches. Most of all, the signal processing of speech enhancement is an interesting problem that’s relevant to digital communications.

>>>>

In this paper we present an educational demonstration in which an \han{inexpensive}, table top Michelson interferometer serves as an analogue to a gravitational-wave detector. 
Vibrations produced by gravitational-wave signals are simulated with sound recordings played though a speaker fixed to the back of one of the interferometer mirrors. 
We consider quasimonochromatic signals as an analogue to continuous wave signals as well as short segments of speech and music which are (more distant) analogues of unmodelled gravitational-wave bursts, e.g. from supernovae. 
% This serves to demonstrate signal processing techniques.

In Sections~\ref{sec:ifo} and ~\ref{sec:single_tone} we demonstrate the recovery of signals with constant and slowly wandering frequency. 
A webcam sampling at \han{XX}$\,{\rm Hz}$ is used to track the interference pattern and the injected signal is recovered succesfully with the Viterbi algorithm. 
These sections direcly demonstrate the continuous wave search techniques used in gravitational-wave data analysis. 
%Using a webcam to track the interference pattern, we inject a wandering frequency signal into the system with a speaker attached to the back of one of mirrors of the interferometer. We then demonstrate how to recover this signal using the Viterbi algorithm, just as is done in continuous gravitational wave searches.

In Section~\ref{sec:complex_audio} we demonstrate recovery of more complex signals of music and speech. 
Substituting the webcam for the increased frequency response of a photodiode read by a Raspberry Pi, we transform the interferometer into an optical microphone.
We demonstrate that the optical microphone is capable of playing back simple recordings well but that it produces unintelligible recordings of speech. 
Mains hum at $50\,{\rm Hz}$ and its harmonics dominates the photodiode output. 
This offers students the opportunity to explore speech enhancement techniques to remove it, including the well performing logMMSE estimator. 
Further avenues of exploration include a more complete model of the optical microphone transfer function, which may lead to improvements in intelligibility.

The demonstration presented here has good potential to be adapted for use in the undergraduate laboratory. 
It can be used to demonstrate topics of interest to physicists and electrical engineers including: interferometer physics; gravitational-wave detection, searches, and analysis; and signal processing with filters and speech enhancement. 
%These demonstrations have good potential to be adapted into the undergraduate laboratory. There is much to explore in the physics of the interferometer and the astronomy of the motivating continuous gravitational wave searches. Most of all, the signal processing of speech enhancement is an interesting problem that’s relevant to digital communications.

### 2020-06-16 paper/ifo-singleTone.tex
going through a few more comments from Andrew

Continuous wave searches look for signals which are close to monochromatic. 
In this section we consider a simple sinusoidal tone; a constant note. 
As described in Section~\ref{sec:ifo}, the audio is played through a speaker fixed to the back of one of the mirrors.

@@

The resulting data is a video capture of the whole interference pattern. 
As only the frequency of the interference ring motion is needed, we take data from a single off-centre pixel indicated by the red dot in Fig.~\ref{fig:interference_pattern}. 
We take the green-channel of the video as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).

As an initial inspection, we play the timeseries of the intensity using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
The two tones sound similar by ear and don't produce any noticeable beating. 
The result is quantified in Fig.~\ref{fig:webcam_spectrum}. 
A tone of $2.09\,{\rm Hz}$ is played through the speaker and the data recorded via webcam. 
Figure~\ref{fig:webcam_spectrum} shows the Fourier transform of the recording.
\han{[to do - rewrite once James has the new numbers (also Fig 3 caption)] 
We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ (where the error represents the $1\sigma$ uncertainty).
The demonstration is able to recover a single frequency tone to within two standard deviations. }
%We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
%The two tones played together sound similar and don’t produce any noticeable beating.
%To quantify the precision, we recorded a signal from an injected tone of $2.09\,{\rm Hz}$. 
%From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$, a result within two standard deviations. 
%The experiment is able to recover a single tone confidently. 
%This showed that the experiment observes a single tone confidently.

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{\label{fig:webcam_spectrum}
Single injected tone results. 
The Fourier amplitude of the intensity timeseres recorded by webcam. 
The frequency of the injection is $2.09\,{\rm Hz}$. 
The recovered frequency peak is at $(2.099\pm 0.008)\,{\rm Hz}$. 
\han{[to do: add new numbers from James]}
%The peak (with amplitude $17.5$) has a frequency was measured at $(2.099\pm 0.008) \mathrm{Hz}$. The actual maximum is achieved at $0\,{\rm Hz}$, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.
}

>>>>

Continuous wave searches look for signals which are close to monochromatic. 
In this section we consider a simple sinusoidal tone; a time at a single constant frequency. 
As described in Section~\ref{sec:ifo}, the audio is played through a speaker fixed to the back of one of the mirrors.

@@

%The resulting data is a video capture of the whole interference pattern. 
%As only the frequency of the interference ring motion is needed, w
We take data from a single off-centre pixel indicated by the red dot in Fig.~\ref{fig:interference_pattern}. 
The green-channel of the video is used as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).
As an initial inspection, we play the timeseries of the intensity using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
The two tones sound similar by ear and don't produce any noticeable beating. 
The result is quantified in Fig.~\ref{fig:webcam_spectrum}. 
A tone of $2.09\,{\rm Hz}$ is played through the speaker and the data recorded via webcam. 
Figure~\ref{fig:webcam_spectrum} shows the Fourier transform of the recording.
\han{[to do - rewrite once James has the new numbers (also Fig 3 caption)] 
We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ (where the error represents the $1\sigma$ uncertainty).
The demonstration is able to recover a single frequency tone to within two standard deviations. }
%We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
%The two tones played together sound similar and don’t produce any noticeable beating.
%To quantify the precision, we recorded a signal from an injected tone of $2.09\,{\rm Hz}$. 
%From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$, a result within two standard deviations. 
%The experiment is able to recover a single tone confidently. 
%This showed that the experiment observes a single tone confidently.

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{\label{fig:webcam_spectrum}
Injected tone at a constant frequency: Fourier amplitude of the intensity time series recorded by the webcam. 
%Single injected tone results. 
%The Fourier amplitude of the intensity timeseres recorded by webcam. 
The frequency of the injection is $2.09\,{\rm Hz}$. 
The recovered frequency peak is at $(2.099\pm 0.008)\,{\rm Hz}$. 
\han{[to do: add new numbers from James]}
%The peak (with amplitude $17.5$) has a frequency was measured at $(2.099\pm 0.008) \mathrm{Hz}$. The actual maximum is achieved at $0\,{\rm Hz}$, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.
}
